Visual Interrogation of Attention-Based Models for Natural Language Inference and Machine Comprehension

### 2018-06-01 EMNLP-demo-paper/conclusion.tex

\section{Discussion}
In this work, we introduce a visualization library for creating customized environments that allow the user to interrogate the relationship between different parts of the model pipeline via interactive queries.
%
We demonstrate the usability and flexibility of tool by configuring the visual components to investigate models for different NLP tasks (e.g., NLI, MC).
%
We also conducted a small-scale user evaluation, where five Ph.D. level students with NLP background spent 30 minutes with the tool and then provided feedback. Most suggested the environment allow them to refine queries iteratively and identify potential issues with the model, but some also mentioned the tool might not provide enough guidance for users do not have an in-depth understanding of the model at hand.

Even though we design the individual components with versatility in mind, due to a large number of variants of attention networks, it is hard to ensure compatibility with all the available configurations.
%
In the future, we plan to improve upon the current attention interface, release the library as an open source package, and expand the visualization components to handle tasks such neural machine translation and more.

>>>>

\section{Discussion}
In this work, we introduce a visualization library for creating customized environments that allow the user to interrogate the relationship between different parts of the model pipeline via interactive queries.
%
We demonstrate the usability and flexibility of the tool by configuring the visual components to investigate models for different NLP tasks (e.g., NLI, MC).
%
We also conducted a small-scale user evaluation, in which five Ph.D. level students with NLP background spent 30 minutes with the tool and then provided feedback. Most suggested the environment allowed them to refine queries iteratively and identify potential issues with the model, but some also mentioned the tool might not provide enough guidance for users who do not have an in-depth understanding of the model at hand.

Even though we designed the individual components with versatility in mind, due to a large number of variants of attention networks, it is hard to ensure compatibility with all the available configurations.
%
In the future, we plan to improve upon the current attention interface, release the library as an open-source package, and expand the visualization components to handle tasks such neural machine translation and more.

### 2018-06-01 EMNLP-demo-paper/result.tex

\subsection{Natural Language Inference}
\label{sec:NLIexample}
The NLI task predicts the relationship between a premise sentence (P) and a hypothesis sentence (H).
The attention matrix captures the alignment of words between these two sentences.
Here we give an example of how a wrong prediction can be corrected by editing the attention value.

@@

As illustrated in Fig.~\ref{fig:NLIexample}(a), the input sentence pair (P:``A woman in a green jacket is drinking tea.'' H:``A woman is drinking green tea.'') is predicted to be \emph{entailment}, which is incorrect.
By examining the attention, we can see the word \emph{green} in ``\emph{green} jacket" is aligned to the \emph{green} in ``\emph{green} tea". However, these two \emph{greens} decorate different nouns, which potentially leads to the wrong prediction. The grammar structure is visually presented by a dependency tree.
%i.e., \emph{green} in \textbf{P} is attached to ``jacket''.
However, the model does not have access to the grammar information and mistakenly assumes the two \emph{greens} are describing the same thing.
% (thus predict \emph{entailment}). 
%
To correct the mistake, we can edit the attention value and remove the align between these two "greens" (see (c)(b)).
As expected, the prediction label is corrected (\emph{neutral}).

\subsection{Machine Comprehension}
\label{sec:MCexample}
In the machine comprehension task, the goal is to select a span of text as the answer
to a question sentence.
%two sequences of texts are given: context and question.
%The goal is to select a span of text from the context that answers the question. 
%We provide a simple model formulation and refer the reader to the paper for details.
%%\begin{align}
%	$C^\prime, Q^\prime = BiLSTM(C), BiLSTM(Q)$,
%	$\overleftarrow{A} = u(C^\prime, Q^\prime)$,
%	$\overrightarrow{A} = v(C^\prime, Q^\prime)$,
%	$s = m(C^\prime, Q^\prime, \overleftarrow{A}, \overrightarrow{A})$,
%	$e = n(s, C^\prime, Q^\prime, \overleftarrow{A}, \overrightarrow{A})$,
%%\end{align}
%where $C$ and $Q$ are embedding matrices for the context and the question,
%$\overleftarrow{A}$ and $\overrightarrow{A}$ are bidirectional attention flows,
%$s$ and $e$ are probabilities for starting and ending indices of the answer span.
%Our system reveals the internal states of $\overleftarrow{A}$, $\overrightarrow{A}$,
%$s$ and $e$.
The attention information is encoded as a bi-directional alignment (i.e., from context to question and vice versa). 
%The key difference is how the question to context alignment is computed (see \cite{Seo2016} for more details).
Here, we refer the context to question attention as \emph{att1}; and the question to context attention as \emph{att2} (which is a vector instead of a matrix according to \citet{Seo2016}). In this demo, we apply a min max normalization for \emph{att2} after the softmax layer to better distinguish different attention values).

>>>>

\subsection{Natural Language Inference}
\label{sec:NLIexample}
The NLI task predicts the entailment relationship between a premise sentence (P) and a hypothesis sentence (H).
The attention matrix captures the alignment of words between these two sentences.
Here we give an example of how a wrong prediction can be corrected by editing attention values.

@@

As illustrated in Fig.~\ref{fig:NLIexample}(a), the input sentence pair (P:``A woman in a green jacket is drinking tea.'' H:``A woman is drinking green tea.'') is predicted to be \emph{entailment}, which is incorrect.
By examining the attention, we can see the word \emph{green} in ``\emph{green} jacket" is aligned to the \emph{green} in ``\emph{green} tea". However, these two \emph{greens} modify different nouns, which potentially leads to the wrong prediction. The grammatical structure is visually shown in the form of the dependency tree.
%i.e., \emph{green} in \textbf{P} is attached to ``jacket''.
However, the model does not have access to the syntactic information and mistakenly assumes the two \emph{greens} modify the same thing.
% (thus predict \emph{entailment}). 
%
To correct the mistake, we can edit the attention value and remove the align between these two "greens" (see (c)(b)).
As expected, the prediction label is corrected (\emph{neutral}).

\subsection{Machine Comprehension}
\label{sec:MCexample}
In the machine comprehension task, the goal is to select a span of text as the answer
to a question sentence.
%two sequences of texts are given: context and question.
%The goal is to select a span of text from the context that answers the question. 
%We provide a simple model formulation and refer the reader to the paper for details.
%%\begin{align}
%	$C^\prime, Q^\prime = BiLSTM(C), BiLSTM(Q)$,
%	$\overleftarrow{A} = u(C^\prime, Q^\prime)$,
%	$\overrightarrow{A} = v(C^\prime, Q^\prime)$,
%	$s = m(C^\prime, Q^\prime, \overleftarrow{A}, \overrightarrow{A})$,
%	$e = n(s, C^\prime, Q^\prime, \overleftarrow{A}, \overrightarrow{A})$,
%%\end{align}
%where $C$ and $Q$ are embedding matrices for the context and the question,
%$\overleftarrow{A}$ and $\overrightarrow{A}$ are bidirectional attention flows,
%$s$ and $e$ are probabilities for starting and ending indices of the answer span.
%Our system reveals the internal states of $\overleftarrow{A}$, $\overrightarrow{A}$,
%$s$ and $e$.
The attention information is encoded as a bi-directional alignment (i.e., from context to question and vice versa). 
%The key difference is how the question to context alignment is computed (see \cite{Seo2016} for more details).
Here, we refer the context to question attention as \emph{att1}; and the question to context attention as \emph{att2} (which is a vector instead of a matrix according to \citet{Seo2016}). In this demo, we apply a min max normalization for \emph{att2} after the softmax layer to better distinguish different attention values.

### 2018-06-01 EMNLP-demo-paper/system.tex

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.9\linewidth]{sentence}
 \vspace{-2mm}
 \caption{
The interface for showing input sentences. The user can manually edit the words or apply automatic perturbation/paraphrasing of the inputs. In the ``perturbed'' drop-down menu, the blue background highlight words not in the original sentence.
 }
 \vspace{-1mm}
\label{fig:sentence}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.8\linewidth]{prediction}
 \caption{
Summarize the prediction results of the perturbed input for the natural language inference model.
the prediction is encoded as a point in the barycentric coordinate system of the triangle, in which each vertex corresponds to one prediction label.
}
\vspace{-3mm}
\label{fig:prediction}
\end{figure}

\subsection{Prediction Visualization}
Efficient visual encoding for prediction is crucial for communicating the model behavior. And to fully support the input perturbation feature, the visual encoding should also allow multiple predictions to be shown in the same visualization.
%
For the natural language inference task, as illustrated in Figure~\ref{fig:prediction}, the predicted probabilities are encoded via a triangular barycentric coordinate system, where the vertices represent the three possible predictions (namely, \emph{entailment}, \emph{contradiction}, and \emph{neutral}). The prediction for the original unperturbed input is illustrated by a larger yellow circle, whereas the prediction of perturbed inputs is represented by smaller grey circles.
A density contour of the prediction is computed to emphasize the highly cluttered areas and detect outliers.

\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=0.95\linewidth]{attentionPanels}
  \vspace{-3mm}
 \caption{
Attention visualization. A bipartite graph encoding is adopted in the graph attention view (a), in which the edge thickness corresponds to the attention value. The same attention values can also be directly visualized in the matrix form (b).
The user can edit the attention values via the pop-up interface illustrated in (c).
We overlay the dependency tree ($a_1$) grammar structure to highlight important words and allow simplification of complex sentences (shown in the video).
%
For highly asymmetric attention, we utilized a zoomable hierarchical visual representation (d). The user can focus on the individual sentence by selecting in the summary visualization.
}
\vspace{-3mm}
\label{fig:attentionVis}
\end{figure*}

@@

We augment these standard visual encoding with grammar structure to address the challenge of long sentences.
The text can be dynamically simplified based on the dependency tree (see Figure~\ref{fig:attentionVis}(a1)). We also show how the dependency information can potentially improve the prediction result in Section~\ref{sec:NLIexample}.
%
To facilitate the perturbation of attention (see Figure~\ref{fig:modelPipeline}), as illustrated in Figure~\ref{fig:attentionVis}(c), we implement an interface for directly manipulating the attention value.

However, when the text sequence becomes significantly longer, i.e., a full paragraph in the machine comprehension task,  even the simplified sentence cannot be meaningfully represented in the graph or matrix visual encoding. To address this visualization challenge, as illustrated in Figure~\ref{fig:attentionVis}(d), we introduce a hierarchical representation. Here, a pure graphical encoding (the color bars marked by \emph{att2}) is used to capture the aggregated attention information for the whole paragraph. The user can focus on localized attention information by selecting a pixel bar that presents a single sentence (the colored blocks in the bar correspond to individual works). We also link this attention representation with matrix form, such that whenever a sentence is selected the local attention is shown in the matrix view (see Figure~\ref{fig:MCexample}(a)).

\subsection{Implementation}
The initial setup cost and unnecessary learning curve are often the barriers to the broad adaptation of a tool. 
Therefore, instead of design visualization system as a monolithic standalone application, 
we implement the proposed system as a python library with modularity and ease of use in mind.
The different pieces of the visualization (i.e., matrix-based attention encoding) can be accessed individually, 
or combined with other components to fit one's workflow via a simple API.
More importantly, the library-based design allows easy integration with the existing model implemented in python.
%To create a visualization, users only need to import the library, create an instance of the visualization object, and specify a set of callback functions, such as generating a prediction, accessing attention, to link the visualization to their NLP models. 
The code for creating an interactive exploration environment for a machine comprehension model is illustrated below.

>>>>

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.9\linewidth]{sentence}
 \vspace{-2mm}
 \caption{
The interface for showing input sentences. The user can manually edit the words or apply automatic perturbation/paraphrasing of the inputs. In the ``perturbed'' drop-down menu, the blue background highlights words not in the original sentence.
 }
 \vspace{-1mm}
\label{fig:sentence}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.8\linewidth]{prediction}
 \caption{
Summary of the prediction results of the perturbed input for the natural language inference model.
The prediction is encoded as a point in the barycentric coordinate system of the triangle, in which each vertex corresponds to one prediction label.
}
\vspace{-3mm}
\label{fig:prediction}
\end{figure}

\subsection{Prediction Visualization}
Efficient visual encoding for prediction is crucial for communicating the model behavior, and to fully support the input perturbation feature, the visual encoding should also allow multiple predictions to be shown in the same visualization.
%

For the natural language inference task, as illustrated in Figure~\ref{fig:prediction}, the predicted probabilities are encoded via a triangular barycentric coordinate system, where the vertices represent the three possible predictions (namely, \emph{entailment}, \emph{contradiction}, and \emph{neutral}). The prediction for the original unperturbed input is illustrated by a larger yellow circle, whereas the prediction of perturbed inputs is represented by smaller grey circles.

A density contour of the prediction is computed to emphasize the highly cluttered areas and detect outliers.

\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=0.95\linewidth]{attentionPanels}
  \vspace{-3mm}
 \caption{
Attention visualization. A bipartite graph encoding is adopted in the graph attention view (a), in which the edge thickness corresponds to the attention value. The same attention values can also be directly visualized in the matrix form (b).
The user can edit the attention values via the pop-up interface illustrated in (c).
We overlay the dependency tree ($a_1$) grammar structure to highlight important words and allow simplification of complex sentences (shown in the video).
%
For highly asymmetric attention, we utilized a zoomable hierarchical visual representation (d). The user can focus on the individual sentence by selecting the summary visualization.
}
\vspace{-3mm}
\label{fig:attentionVis}
\end{figure*}

@@

We augment these standard visual encodings with grammar structure to address the challenge of long sentences.
The text can be dynamically simplified based on the dependency tree (see Figure~\ref{fig:attentionVis}(a1)). We also show how the dependency information can potentially improve the prediction result in Section~\ref{sec:NLIexample}.
%
To facilitate the perturbation of attention (see Figure~\ref{fig:modelPipeline}), as illustrated in Figure~\ref{fig:attentionVis}(c), we implement an interface for directly manipulating the attention value.

However, when the text sequence becomes significantly longer, i.e., a full paragraph in the machine comprehension task,  even the simplified sentence cannot be meaningfully represented in the graph or matrix visual encoding. To address this visualization challenge, as illustrated in Figure~\ref{fig:attentionVis}(d), we introduce a hierarchical representation. Here, a pure graphical encoding (the color bars marked by \emph{att2}) is used to capture the aggregated attention information for the whole paragraph. The user can focus on localized attention information by selecting a pixel bar that represents a single sentence (the colored blocks in the bar correspond to individual works). We also link this attention representation with the matrix form, such that whenever a sentence is selected the local attention is shown in the matrix view (see Figure~\ref{fig:MCexample}(a)).

\subsection{Implementation}
The initial setup cost and unnecessary learning curve are often the barriers to broad adaptation of a tool. 
Therefore, instead of designing a visualization system as a monolithic standalone application, 
we implement the proposed system as a Python library with modularity and ease of use in mind.
The different pieces of the visualization (i.e., matrix-based attention encoding) can be accessed individually, 
or combined with other components to fit one's work-flow via a simple API.
More importantly, the library-based design allows easy integration with the existing model implemented in python.
%To create a visualization, users only need to import the library, create an instance of the visualization object, and specify a set of callback functions, such as generating a prediction, accessing attention, to link the visualization to their NLP models. 
The code for creating an interactive exploration environment for a machine comprehension model is illustrated below.
